---
title: 'powerLATE: Generalized Power Analysis for LATE'
output:
  html_document: default
  pdf_document: default
---

***

# Examples

Suppose we want to recover power, without making the ordered-means assumption, 

we are interested in the power of an experiment, in which exactly half of the experimental subjects will be assigned to treatment and the other half to control. Further suppose the sample size is 800, the compliance rate is 0.35 and the treatment effect size is 0.8. We can compute the power of this experiment by the following:

```{r}
library(powerLATE)
res <- powerLATE(pZ = 0.5, pi = 0.35, N = 800, kappa = 0.8)
```

In general, two parameters from the set {effect size (`kappa`), sample size (`N`), power (`power`)} must be specified, from which the third (target) parameter will be calculated. For example, if we are interested in the lower bound on the required sample size for a level of power of 0.8 to reject the null hypothesis, we can run the following:

```{r}
res <- powerLATE(pZ = 0.5, pi = 0.35, kappa = 0.8, power = 0.8)
```

If absolute effect is desired instead of effect size, simply set `effect.size=FALSE` and specify `tau`, the hypothesized absolute effect, and `omega`, the within-group standard deviation of the outcome.
```{r}
res <- powerLATE(pZ = 0.5, pi = 0.35, N = 800, effect.size = FALSE, tau = 0.4, omega = 0.6,)
```

We can also input a vector of values for any one parameter in the set \{pi, N, kappa, power, tau, omega\}.

```{r}
res <- powerLATE(pZ = 0.5, pi = 0.35, N = 800, kappa = seq(0.5, 1.0, 0.1))
```

With covariates, use `powerLATE.cov` and specify `r2dw` and `r2yw`.

```{r}
res <- powerLATE.cov(pZ = 0.5, pi = 0.35, N = 800, kappa = seq(0.5, 1.0, 0.1), r2dw = 0.2, r2yw = 0.3)
```

In addition to the message that is automatically printed out, `powerLATE` and `powerLATE.cov` also return the output parameter values as an invisible object, simply query

```{r}
res$output.parameter
```

For more details and examples, see `?powerLATE` and `?powerLATE.cov`

***