---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# powerLATE

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/EddieYang211/powerLATE.svg?branch=master)](https://travis-ci.com/EddieYang211/powerLATE)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/EddieYang211/powerLATE?branch=master&svg=true)](https://ci.appveyor.com/project/EddieYang211/powerLATE)
[![Codecov test coverage](https://codecov.io/gh/EddieYang211/powerLATE/branch/master/graph/badge.svg)](https://codecov.io/gh/EddieYang211/powerLATE?branch=master)
<!-- badges: end -->

The goal of powerLATE is to provide a generalized approach power analysis in the context of estimating a local average treatment effect (LATE), where the study subjects exhibit noncompliance with treatment assignment. powerLATE uses standardized effect sizes to place bounds on the power for the most commonly used estimator of the LATE, the Wald IV estimator, whereby variance terms and distributional parameters need not be specified nor assumed. Instead, in addition to the effect size, sample size, and error tolerance parameters, the only other parameter that must be specified by the researcher is the compliance rate. Additional conditions can also be introduced to further narrow the bounds on the power calculation.

## Installation

You can install the released version of powerLATE from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("powerLATE")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(powerLATE)
## basic example code
res <- powerLATE.cov(pZ=0.5, pi=0.35, kappa=seq(0.4, 1.0, 0.1), power=0.8,
                     r2dw=0.2, r2yw=0.15)

# also returns an invisible object: output.parameter
res$output.parameter

res <- powerLATE.cov(pZ=0.67, pi=0.35, kappa=seq(0.4, 1.0, 0.1), 
                power=0.8, assume.ord.means=TRUE, r2dw=0.2, r2yw=0.15)
```
