---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# powerLATE

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/EddieYang211/powerLATE.svg?branch=master)](https://travis-ci.com/EddieYang211/powerLATE)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/EddieYang211/powerLATE?branch=master&svg=true)](https://ci.appveyor.com/project/EddieYang211/powerLATE)
[![Codecov test coverage](https://codecov.io/gh/EddieYang211/powerLATE/branch/master/graph/badge.svg)](https://codecov.io/gh/EddieYang211/powerLATE?branch=master)
<!-- badges: end -->

powerLATE implements the generalized power analysis for the local average treatment effect (LATE), proposed by [Bansak (2020)](https://projecteuclid.org/download/pdfview_1/euclid.ss/1591171230).

Power analysis is in the context of estimating the LATE (also known as the complier average causal effect, or CACE), with calculations based on a test of the null hypothesis that the LATE equals 0 with a two-sided alternative. The method uses standardized effect sizes to place a conservative bound on the power under minimal assumptions. powerLATE allows users to recover power, sample size requirements, or minimum detectable effect sizes. It also allows users to work with absolute effects rather than effect sizes, to specify an additional assumption to narrow the bounds, and to incorporate covariate adjustment.

## Installation

You can install the released version of powerLATE from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("powerLATE")
```
Or the development version from GitHub:

``` r
# install.packages("devtools")
devtools::install_github("EddieYang211/powerLATE")
```

## Getting started

```{r}
library(powerLATE)
```

powerLATE provides two main functions:

* `powerLATE`, which computes the power of the Wald IV estimator, or determines parameters to obtain a target power.

* `powerLATE.cov`, which is similar to `powerLATE` but additioanlly allows the inclusion of covariates. 


## Examples

Suppose we are interested in the power of an experiment, in which exactly half of the experimental subjects will be assigned to treatment and the other half to control. Further suppose the sample size is 800, the compliance rate is 0.35 and the treatment effect size is 0.8. We can compute the power of this experiment by the following:

```{r}
library(powerLATE)
res <- powerLATE(pZ = 0.5, pi = 0.35, N = 800, kappa = 0.8)
```

In general, two parameters from the set {effect size (`kappa`), sample size (`N`), power (`power`)} must be specified, from which the third (target) parameter will be calculated. For example, if we are interested in the lower bound on the required sample size for a level of power of 0.8 to reject the null hypothesis, we can run the following:

```{r}
res <- powerLATE(pZ = 0.5, pi = 0.35, kappa = 0.8, power = 0.8)
```

If absolute effect is desired instead of effect size, simply set `effect.size=FALSE` and specify `tau`, the hypothesized absolute effect, and `omega`, the within-group standard deviation of the outcome.
```{r}
res <- powerLATE(pZ = 0.5, pi = 0.35, N = 800, effect.size = FALSE, tau = 0.4, omega = 0.6,)
```

We can also input a vector of values for any one parameter in the set \{pi, N, kappa, power, tau, omega\}.

```{r}
res <- powerLATE(pZ = 0.5, pi = 0.35, N = 800, kappa = seq(0.5, 1.0, 0.1))
```

With covariates, use `powerLATE.cov` and specify `r2dw` and `r2yw`.

```{r}
res <- powerLATE.cov(pZ = 0.5, pi = 0.35, N = 800, kappa = seq(0.5, 1.0, 0.1), r2dw = 0.2, r2yw = 0.3)
```

In addition to the message that is automatically printed out, `powerLATE` and `powerLATE.cov` also return the output parameter values as an invisible object, simply query

```{r}
res$output.parameter
```

For more details and examples, see `?powerLATE` and `?powerLATE.cov`

## Reference

For a detailed description of the method see:

- [Bansak, K. (2020). A Generalized Approach to Power Analysis for Local Average Treatment Effects](https://projecteuclid.org/download/pdfview_1/euclid.ss/1591171230)

## Maintainer 

- [Eddie Yang](https://github.com/EddieYang211)
